---
title: "Week3A.GlobalB"
author: "Caresse Cross Beard"
format: html
editor: visual
---

## Approach

Connect to my database in pgadmin4 and load the movies data I created. I will calculate the global baseline for a movie recommendation.

# Code

```{r load packages}
library(dplyr)
library(DBI) 
library(RPostgres) 
```

```{r connect to database}
con <- dbConnect(
  RPostgres::Postgres(),
  dbname = "MoviesCCB",
  host = "localhost",
  port = 5432,
  user = "postgres",
  password = Sys.getenv("DB_PASSWORD")
)
```

```{r get table}
ratings <- dbGetQuery(con, "
  SELECT *
  FROM ratings
")

```

```{r check table}
head(ratings)
```

```{r}
ratings <- dbReadTable(con, "ratings")
users <- dbReadTable(con, "users")
movies <- dbReadTable(con, "movies")


```

```{r disconnect}
dbDisconnect(con)
```

```{r calculate overall average}
global_mean <- mean(ratings$rating_value, na.rm = TRUE)
global_mean
```

The average movie rating was 3.58. Next, I will find each persons bias. Using the mean of the rating value and subtract the global mean.

```{r calculate user bias}
user_bias <- ratings %>%
  group_by(user_id) %>%
  summarise(
    user_mean = mean(rating_value, na.rm = TRUE),
    user_bias = user_mean - global_mean
  )
user_bias
```

```{r check names}
user_bias_named <- user_bias %>%
  left_join(users, by = "user_id") %>%
  select(user_name, user_mean, user_bias)
user_bias_named
```

My friend Jennifer tended to rate higher than the other friends. Then, I will check the movie rating average.

```{r calculate movie bias}
movie_bias <- ratings %>%
  group_by(movie_id) %>%
  summarise(
    movie_mean = mean(rating_value, na.rm = TRUE),
    movie_bias = movie_mean - global_mean
  )
movie_bias
```

```{r show movie bias titles}
movies_named <- movie_bias %>%
  left_join(movies, by = "movie_id") %>%
  select(movie_title, movie_mean, movie_bias)
movies_named
```

Among my friends, Spider-man: Brand New Day was generally rated higher than the others.

```{r}
baseline_predictions <- ratings %>%
  select(user_id, movie_id) %>%
  distinct() %>%
  left_join(user_bias, by = "user_id") %>%
  left_join(movie_bias, by = "movie_id") %>%
  mutate(
    predicted_rating = global_mean + user_bias + movie_bias
  )
baseline_predictions
```

```{r}
baseline_predictions_named <- baseline_predictions %>%
  left_join(users, by = "user_id") %>%
  left_join(movies, by = "movie_id") %>%
  select(user_name, movie_title, predicted_rating)
```

```{r check Jennifer}
baseline_predictions_named %>%
  filter(user_name == "Jennifer") %>%
  arrange(desc(predicted_rating))

```
