---
title: "MoviesCCB2A"
author: "Caresse Cross Beard"
format: html
editor: visual
---

## Approach

Install pgAdmin4 and create database to analyse movies and their ratings.

## PgAdmin4

I inputted data into my database I created in pgAdmin4. I inserted tables for user, movies and ratings.

```{r}
library(DBI) 
library(RPostgres) 
library(dplyr) 
library(tidyr)
```

## Connect to the database

I connect to the database below. I figured out how to add it and hide my password. I used Gemini to help me figure out how to do that in Rstudio.

```{r}
con <- dbConnect(
  RPostgres::Postgres(),
  dbname = "MoviesCCB",
  host = "localhost",
  port = 5432,
  user = "postgres",
  password = Sys.getenv("DB_PASSWORD")
)

```

## Fetch data frames

```{r}
ratings <- dbGetQuery(con, "
  SELECT
    u.user_name,
    m.movie_title,
    r.rating_value
  FROM Ratings r
  JOIN Users u
    ON r.user_id = u.user_id
  JOIN Movies m
    ON r.movie_id = m.movie_id;
")
```

```{r}
users <- dbReadTable(con, "users")
movies <- dbReadTable(con, "movies")
ratings <- dbReadTable(con, "ratings")
```

```{r check tables}
dbListTables(con)
```

```{r disconnect from database}
dbDisconnect(con)
```

```{r list data from tables}
users
movies
ratings
```

```{r join three tables together}
movie_ratings <- ratings %>%
  left_join(users, by = "user_id") %>%
  left_join(movies, by = "movie_id")
```

```{r see new table}
movie_ratings
```

```{r add characters to table}
complete_ratings <- movie_ratings %>%
  complete(
    user_name,
    movie_title
  )
```

```{r show NA ratings}
complete_ratings %>%
  filter(is.na(rating_value))
```

## Movie rating summary

```{r summarize by movie}
movie_summary <- complete_ratings %>%
  group_by(movie_title) %>%
  summarise(
    average_score = mean(rating_value, na.rm = TRUE),
    missing_ratings = sum(is.na(rating_value)),
    .groups = "drop"
  )

```

```{r}
movie_summary
```

## Summary by user

```{r summarize by user}
user_summary <- complete_ratings %>%
  group_by(user_name) %>%
  summarise(
    movies_watched = sum(!is.na(rating_value)),
    average_score = mean(rating_value, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
user_summary
```

## Conclusion

From my data, the movie with the highest average rating was *Spiderman: Brand New Day*. The second highest average rated movie is a tie between *The Drama* and *Odyssey.* Spiderman was the highest rated and all 5 friends had seen the movie.
