---
title: "Penguins_Classification_Metrics_CCB"
Date: "09/10/2026"
format: html
---

# Penguin Predictions Approach

In order to practice classification metrics, I will first look over the data and then run a series of codes to test it's accuracy.

As I work with the data to calculate the null rate and experiment with different probability thresholds.

### Null Rate Error

From the penguins predictions, we can see that there are 54 males and 39 females. Classifying the majority of the penguins male would mean we would have a null rate error of 41.9%. Males are the majority in this dataset.

```{r load date} #loading data}
library(tidyverse) 
penguin_predictions <- read_csv("penguin_predictions.csv")  
```

```{r glimpse}
glimpse(penguin_predictions)
```

```{r } #count of each class}
table(penguin_predictions$sex)
```

```{r null error rate} #calcute null error rate}
gender_counts <-table(penguin_predictions$sex)  
null_error_rate <- 1 - max(gender_counts)/sum(gender_counts)  
null_error_rate
```

### Confusion Matrix

We can measure the thresholds. The tables below count the true positives, false positives, true negatives, and false negatives produced using the 0.2, 0.5, and 0.8 threshold.

```{r}
penguin_predictions <- penguin_predictions %>%   
  mutate(     
    pred_2 = if_else(.pred_female >= 0.2, "female", "male"),     
    pred_5 = if_else(.pred_female >= 0.5, "female", "male"),     
    pred_8 = if_else(.pred_female >= 0.8, "female", "male")  
    ) 
penguin_predictions 
```

```{r}
table(   
  Predicted = penguin_predictions$pred_2,   
  Actual = penguin_predictions$sex ) 
```

```{r}
table(   
  Predicted = penguin_predictions$pred_5,   
  Actual = penguin_predictions$sex )
```

```{r}
table(   Predicted = penguin_predictions$pred_8,   
         Actual = penguin_predictions$sex )
```

### Explanation

For each threshold, the predicted sex will be female. If 20% chance the penguin is female, then call it female otherwise it is male. In the first code chunk, we see 32 females as TP, 6 actual female categorized as male meaning FP, 2 actual male categorized as female meaning FN, and 48 actual male categorized as male as TN.

### Performance Metrics

Accuracy = How often was the prediction correct overall? Precision = Of the penguins predicted to be female, how many actually were female? \
Recall = Of the penguins who were actually female, how many did the model identify? \
F1 score = A combined measure of precision and recall.

```{r}
```
